---
title: "Shape Bias Benchmark — Results"
format:
  html:
    embed-resources: true
---

```{r setup, include=FALSE}
source("src/load_data.R")
source("src/plot_model_bias.R")
source("src/plot_stimulus_bias.R")
source("src/plot_position_bias.R")
source("src/plot_word_sensitivity.R")

df <- load_results(get_data_path()) |> add_model_size()
fig_dir <- get_figures_dir()
```

## Validation

Before interpreting shape-bias results, we verify that models are performing the task validly. Three checks: (1) position bias — does the model track images or just pick a fixed position? (2) word sensitivity — does the model's choice change with different words? (3) word type effects — do pseudo-words vs random strings produce different patterns?

### Position Bias

For each model, we check whether responses are driven by image content or by position bias. For every stimulus-word pair run in both orderings (shape_first and texture_first), we test whether the model's answer follows the correct image when positions are swapped. A model that tracks the image should pick "1" when its preferred match is first and "2" when it is second.

```{r fig-position-bias, fig.width=8, fig.height=5, fig.cap="Proportion of paired trials where the model tracks the correct image vs shows position bias. Models below 50% are not reliably performing the task."}
p <- plot_position_bias_main(df)
ggsave(file.path(fig_dir, "C_position_bias_validation.png"), p, width = 8, height = 5, dpi = 300)
p
```

```{r fig-position-bias-stim, fig.width=10, fig.height=5, fig.cap="Image tracking rate by stimulus, broken down by model."}
p <- plot_position_bias_by_stimulus(df)
ggsave(file.path(fig_dir, "C_position_bias_by_stimulus.png"), p, width = 10, height = 5, dpi = 300)
p
```

```{r fig-position-bias-ranked, fig.width=10, fig.height=5, fig.cap="Rank-ordered image tracking rate by model. Each model's stimuli sorted independently by tracking rate. Labels show stimulus ID."}
p <- plot_position_bias_ranked(df)
ggsave(file.path(fig_dir, "C_position_bias_ranked.png"), p, width = 10, height = 5, dpi = 300)
p
```

### Word Sensitivity

For each (model, stimulus, ordering) group, we check whether the model gives the same choice for all words or whether different words lead to different choices. A model that always picks the same option regardless of word is "word-insensitive" — suggesting it ignores the word entirely.

```{r fig-word-sensitivity, fig.width=8, fig.height=5, fig.cap="Proportion of stimulus-ordering groups where the model's choice varies across words."}
p <- plot_word_sensitivity_main(df)
ggsave(file.path(fig_dir, "D_word_sensitivity_validation.png"), p, width = 8, height = 5, dpi = 300)
p
```

```{r fig-word-sensitivity-stim, fig.width=10, fig.height=5, fig.cap="Word sensitivity rate by stimulus, broken down by model."}
p <- plot_word_sensitivity_by_stimulus(df)
ggsave(file.path(fig_dir, "D_word_sensitivity_by_stimulus.png"), p, width = 10, height = 5, dpi = 300)
p
```

### Word Type Effects

We compare shape-bias rates between pseudo-words (word-like strings) and random strings. If a model responds differently to these two word types, it suggests linguistic processing influences its visual judgments.

```{r fig-word-type, fig.width=8, fig.height=5, fig.cap="Mean shape-bias by word type (pseudo-words vs random strings) per model, with SE error bars."}
p <- plot_word_type_effect(df)
ggsave(file.path(fig_dir, "D_word_type_effect.png"), p, width = 8, height = 5, dpi = 300)
p
```

### Shape Bias by Word

We break down shape bias for each individual word, faceted by word type (pseudo-words vs random strings). Words are ordered by length to reflect the length-matched pairs design.

```{r fig-word-bias, fig.width=10, fig.height=5, fig.cap="Shape bias for each word, faceted by word type. Words ordered by length. Dashed line = chance (50%)."}
p <- plot_word_bias(df)
ggsave(file.path(fig_dir, "D_word_bias_by_word.png"), p, width = 10, height = 5, dpi = 300)
p
```

```{r fig-word-bias-by-model, fig.width=10, fig.height=5, fig.cap="Shape bias for each word broken down by model. Words ordered by length. Dashed line = chance (50%)."}
p <- plot_word_bias_by_model(df)
ggsave(file.path(fig_dir, "D_word_bias_by_word_by_model.png"), p, width = 10, height = 5, dpi = 300)
p
```

## Shape Bias by Model

```{r fig-model-main, fig.width=8, fig.height=5, fig.cap="Proportion of shape choices by model, sorted by parameter count. Dashed line = chance (0.5)."}
p <- plot_model_bias_main(df)
ggsave(file.path(fig_dir, "A_model_bias_main.png"), p, width = 8, height = 5, dpi = 300)
p
```

```{r fig-model-supplement, fig.width=8, fig.height=5, fig.cap="Full response breakdown (shape / texture / unclear) by model."}
p <- plot_model_bias_supplement(df)
ggsave(file.path(fig_dir, "A_model_bias_supplement.png"), p, width = 8, height = 5, dpi = 300)
p
```

### By order method

```{r fig-model-deterministic, fig.width=8, fig.height=5, fig.cap="Shape bias — deterministic trials only (both orderings run per stimulus)."}
p <- plot_model_bias_main(df, order_method = "deterministic")
ggsave(file.path(fig_dir, "A_model_bias_deterministic.png"), p, width = 8, height = 5, dpi = 300)
p
```

```{r fig-model-random-method, fig.width=8, fig.height=5, fig.cap="Shape bias — random ordering trials only (one random ordering per stimulus)."}
p <- plot_model_bias_main(df, order_method = "random")
ggsave(file.path(fig_dir, "A_model_bias_random_ordering.png"), p, width = 8, height = 5, dpi = 300)
p
```

### By ordering

```{r fig-model-shape-first, fig.width=8, fig.height=5, fig.cap="Shape bias — shape_first ordering only."}
p <- plot_model_bias_main(df, ordering = "shape_first")
ggsave(file.path(fig_dir, "A_model_bias_shape_first.png"), p, width = 8, height = 5, dpi = 300)
p
```

```{r fig-model-texture-first, fig.width=8, fig.height=5, fig.cap="Shape bias — texture_first ordering only."}
p <- plot_model_bias_main(df, ordering = "texture_first")
ggsave(file.path(fig_dir, "A_model_bias_texture_first.png"), p, width = 8, height = 5, dpi = 300)
p
```

### By word type

```{r fig-model-sudo, fig.width=8, fig.height=5, fig.cap="Shape bias — pseudo-words only."}
p <- plot_model_bias_main(df, word_type = "sudo")
ggsave(file.path(fig_dir, "A_model_bias_sudo.png"), p, width = 8, height = 5, dpi = 300)
p
```

```{r fig-model-random, fig.width=8, fig.height=5, fig.cap="Shape bias — random strings only."}
p <- plot_model_bias_main(df, word_type = "random")
ggsave(file.path(fig_dir, "A_model_bias_random.png"), p, width = 8, height = 5, dpi = 300)
p
```

## Shape Bias by Stimulus

```{r fig-stim-main, fig.width=10, fig.height=5, fig.cap="Proportion of shape choices by stimulus, averaged across models. Dashed line = chance (0.5)."}
p <- plot_stimulus_bias_main(df)
ggsave(file.path(fig_dir, "B_stimulus_bias_main.png"), p, width = 10, height = 5, dpi = 300)
p
```

```{r fig-stim-supplement, fig.width=10, fig.height=5, fig.cap="Full response breakdown (shape / texture / unclear) by stimulus."}
p <- plot_stimulus_bias_supplement(df)
ggsave(file.path(fig_dir, "B_stimulus_bias_supplement.png"), p, width = 10, height = 5, dpi = 300)
p
```

### By order method

```{r fig-stim-deterministic, fig.width=10, fig.height=5, fig.cap="Stimulus bias — deterministic trials only (both orderings run per stimulus)."}
p <- plot_stimulus_bias_main(df, order_method = "deterministic")
ggsave(file.path(fig_dir, "B_stimulus_bias_deterministic.png"), p, width = 10, height = 5, dpi = 300)
p
```

```{r fig-stim-random-method, fig.width=10, fig.height=5, fig.cap="Stimulus bias — random ordering trials only (one random ordering per stimulus)."}
p <- plot_stimulus_bias_main(df, order_method = "random")
ggsave(file.path(fig_dir, "B_stimulus_bias_random_ordering.png"), p, width = 10, height = 5, dpi = 300)
p
```

### By ordering

```{r fig-stim-shape-first, fig.width=10, fig.height=5, fig.cap="Stimulus bias — shape_first ordering only."}
p <- plot_stimulus_bias_main(df, ordering = "shape_first")
ggsave(file.path(fig_dir, "B_stimulus_bias_shape_first.png"), p, width = 10, height = 5, dpi = 300)
p
```

```{r fig-stim-texture-first, fig.width=10, fig.height=5, fig.cap="Stimulus bias — texture_first ordering only."}
p <- plot_stimulus_bias_main(df, ordering = "texture_first")
ggsave(file.path(fig_dir, "B_stimulus_bias_texture_first.png"), p, width = 10, height = 5, dpi = 300)
p
```

### By word type

```{r fig-stim-sudo, fig.width=10, fig.height=5, fig.cap="Stimulus bias — pseudo-words only."}
p <- plot_stimulus_bias_main(df, word_type = "sudo")
ggsave(file.path(fig_dir, "B_stimulus_bias_sudo.png"), p, width = 10, height = 5, dpi = 300)
p
```

```{r fig-stim-random, fig.width=10, fig.height=5, fig.cap="Stimulus bias — random strings only."}
p <- plot_stimulus_bias_main(df, word_type = "random")
ggsave(file.path(fig_dir, "B_stimulus_bias_random.png"), p, width = 10, height = 5, dpi = 300)
p
```
